---
title: "Custom theme for a `ggplot2` graph"
author: "Constantin Manuel Bosancianu"
date: "May 7, 2021"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    code_folding: hide
    highlight: pygments
---

The following lines of code prepare a custom theme for the graphs that will be  produced. If you just add this theme at the end of every plot with `+ theme()` you will get the same design. Feel free to customize this theme as much as you like.

```{r setup-packages, warning=FALSE, message=FALSE, comment=NA, results='hide'}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      comment = NA,
                      message = FALSE)

library(pacman)
p_load(ggplot2)

set.seed(475934)
```

Create 2 correlated variables for a scaterplot.

```{r create-data}
a <- rnorm(100, 0, 1)
b <- 0.2 + 0.45 * a + rnorm(100, 0, 0.5)

df_plot <- data.frame(A = a,
                      B = b)
```

# Initial version

First version of the plot.

```{r first-version, fig.height=4, fig.width=6, fig.align='center'}
ggplot(df_plot,
       aes(x = A,
           y = B)) +
    geom_point(size = 2) +
    labs(x = "A (independent)",
         y = "B (dependent)",
         title = "Example of scatterplot",
         caption = "Note: Simulated data")
```

# Customized version

Customize theme.

```{r customized-version, fig.height=4, fig.width=6, fig.align='center'}
ggplot(df_plot,
       aes(x = A,
           y = B)) +
    geom_point(size = 2) +
    labs(x = "A (independent)",
         y = "B (dependent)",
         title = "Example of scatterplot",
         caption = "Note: Simulated data") +
    theme(plot.title = element_text(size = 18, face = "bold", color = "black"),
          # Controls the main title of the graph: 18 points, bold point, black color.
          axis.title.x = element_text(size = 16, face = "bold", color = "black", angle = 0),
          # Controls the label of the X axis.
          axis.title.y = element_text(size = 16, face = "bold", color = "black", angle = 90),
          # Controls the label of the Y axis.
          axis.text.x = element_text(size = 12, color = "black", angle = 45),
          # Controls the X-axis text (usually, numbers which are plotted on the axis).
          axis.text.y = element_text(size = 12, color = "black", angle = 45, vjust = 0),
          # Controls the Y-axis text.
          panel.grid.major.x = element_blank(),
          # Controls whether the X axis should have thick grid lines.
          # Sometimes, e.g. if you're plotting longitudinal trends,
          # you may not want vertical grid lines.
          # If you nevertheless want vertical grid lines, use
          # "element_line()" instead of "element_blank()", and specify
          # the characteristics of the line (color, size, line type).
          panel.grid.minor.x = element_blank(),
          # Controls whether the X axis should have thin grid lines.
          panel.grid.major.y = element_line(color = "grey80"),
          # Controls whether the Y axis should have thick horizontal lines.
          panel.grid.minor.y = element_line(color = "grey40"),
          # Controls whether the Y axis should have thin horizontal lines.
          panel.background = element_rect(fill = "white", color = NA),
          # Controls the panel background (whether it's grey (the
          # default), or white, or whatever color you need).
          plot.background = element_rect(fill = "white", color = NA),
          # Controls the entire plot background (whether it's white,
          # or light blue (as in Stata)).
          legend.title = element_text(color = "black", face = "bold", size = 12),
          # Controls the title of the legend (not a feature in this
          # specific plot). 
          legend.key = element_rect(color = "white"),
          # Controls the background color of the legend text (not a
          # feature in this specific plot).
          legend.text = element_text(color = "black", size = 10),
          # Controls the text of the legend elements (not a feature in
          # this specific plot).
          legend.position = "right",
          # Controls the position of the legend (not a feature in this
          # specific plot).
          axis.ticks = element_blank())
          # Controls whether there should be small notches on the
          # axes, where the numbers are, or not. I prefer not to have
          # them there.
```

If you want to, you can save this entire theme specification in a new object called `personal_theme`. If you want to apply it to every plot, you would just add `+ personal_theme()` at the end of each plot, and it would be formatted in the same way.
